---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# simplerspec.opus

The goal of simplerspec.opus is to facilitate spectral measurement workflows
for Bruker spectrometers. It is tailored to file outputs and directory 
structures produced by the Bruker OPUS and OPUS Lab software.

## Installation and prerequisites

You can install the most recent version of simplerspec.opus using

``` r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("philipp-baumann/simplerspec.opus")
```

This package assumes that OPUS measurements are organized in folders (one per measurement day). 

## Perform folder, file, and data and metadata integrity tests

This test verifies whether both metadata and data are
complete, for each of the "date" folders separately. The binary Bruker OPUS
files, that contain all spectral measurements and aquisition parameters, are 
located in folders named by date. The names are unluckily in the format 
`"YYYY_M_D"`. That does not follow  
[ISO 8601](https://en.wikipedia.org/wiki/ISO_8601).

```{r, eval = FALSE}
# Load packages
pkgs <- c("simplerspec.opus", "here")
lapply(pkgs, library, character.only = TRUE)
```

You can do the test using the function below:

```{r, eval = FALSE}
test_data_metadata(data_root = here("data", "spectra", "2018-BDM"))
```

This test ensures that there is a minimum chance that samples were wrongly 
labelled or measurement positions were swapped.

   
### Test category 1:

* **Expectation:** Metadata are Open Office spreadsheets with extension
  `.ods`. The file name has a date prefix of format `YYYY-MM-DD`.

* **Test:** Test whether every measurement folder has a corresponding 
  metadata file. This is done by extracting `"YYYY_M_D"` from
  the metadata file names, and converting to `"YYYY-MM-DD"`, and check if
  metadata file name information and and folder names are all identical.
  
### Test category 2:

* **Expectation:** Every metadata spreadsheet has a `date` column with 
  header named identically, where entries represent dates in format
  `YYYY-MM-DD`.

* **Test:** Test whether the `date` column entries are matching the expanded
  measurement folders for each date.
  
### Tests category 3: All OPUS files have the correct complete metadata and vice versa

  
* **Expectations:** The metadata spreadsheets contain a column `pos` that
  specifies the well plate positions, and also a column `sample_id`. The
  `sample_id` of a sample needs to be unique. The OPUS files that are generated 
  after measurements have a `_<plate-position>` suffix after a `sample_id`
  string.
  
* **Test:** In order to check that there is a metadata entry for each measured
  file, file names are reconstructed from `sample_id` and `pos` information. 
  Further, repetition numbers starting from 0 are sequentially numbered by an
  integer increasing by 1, are reconstructed as file extension  
  `.<repetition_number>`. This is what is expected to be generated from the 
  measurements managed by the Bruker OPUS lab interface.
  
  
This could be an example test message when data and metadata are not following
expectations:

```{r, eval = FALSE}
# > test_data_metadata(data_root = here("data", "spectra", "2018-BDM"))
# Each measurement folder has a corresponding metadata file.
# `date` column entries in the metadata spreadsheet(s)
#       correspond to the expected measurement folder name (date) in the metadata
#       file name(s)
# Error in struct_data_metadata(data_root = here("data", "spectra", "2018-BDM")) : 
#   Measurment data files and metadata are not complete. Metadata entries for the following measurement data <files are missing: list(date = c("2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02"), sample_id = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA), sample_id_pos = c("NA_A4", "NA_B4", "NA_C4", "NA_D4", "NA_A5", "NA_B5", "NA_C5", "NA_D5", "NA_A6", "NA_B6", "NA_C6", "NA_D6"), plate_no = c(6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6), pos = c("A4", "B4", "C4", "D4", "A5", "B5", "C5", "D5", "A6", "B6", "C6", "D6"
# ), meas_folder_exp = c("2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02", "2018-10-02"), meas_rep_meta = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), file_id = c("NA_A4.0", "NA_B4.0", "NA_C4.0", "NA_D4.0", "NA_A5.0", "NA_B5.0", "NA_C5.0", "NA_D5.0", "NA_A6.0", "NA_B
```

In that case, have a look at the message(s), check the metadata spreadsheets,
fix accordingly, and finally re-test. If tests succeed, you will receive e.g.

```{r, eval = FALSE}
# > test_data_metadata(data_root = here("data", "spectra", "2018-BDM"))
# Each measurement folder has a corresponding metadata file.
# `date` column entries in the metadata spreadsheet(s)
#       correspond to the expected measurement folder name (date) in the metadata
#       file name(s)
# Measurement data files and metadata records are complete.
# # A tibble: 3,438 x 10
#    date  sample_id sample_id_pos plate_no pos   meas_folder_exp meas_rep_meta file_id
#    <chr> <chr>     <chr>            <int> <chr> <chr>                   <int> <chr>  
#  1 2018… 30017 KB… 30017 KB017 …        1 B1    2018-09-03                  0 30017 …
#  2 2018… 30017 KB… 30017 KB017 …        3 B1    2018-09-03                  1 30017 …
#  3 2018… 30017 KB… 30017 KB017 …        4 B1    2018-09-03                  2 30017 …
#  4 2018… 30017 KB… 30017 KB017 …        5 B1    2018-09-03                  3 30017 …
#  5 2018… 30017 KB… 30017 KB017 …        6 B1    2018-09-03                  4 30017 …
#  6 2018… 30017 KB… 30017 KB017 …        7 B1    2018-09-03                  5 30017 …
#  7 2018… 30017 KB… 30017 KB017 …        8 B1    2018-09-03                  6 30017 …
#  8 2018… 30017 KB… 30017 KB017 …        9 B1    2018-09-03                  7 30017 …
#  9 2018… 30017 KB… 30017 KB017 …        2 D1    2018-09-03                  0 30017 …
# 10 2018… 30015 KB… 30015 KB015 …        1 C1    2018-09-03                  0 30015 …
# # ... with 3,428 more rows, and 2 more variables: sample_id_reps <int>,
# #   date_file <chr>
```

# Read spectra

Henceforth, time is ready to read the spectra ;-)

```{r, eval = FALSE}
future::plan(future::multiprocess)

# Read OPUS test files
files_wizardproj_2019 <-  list_opus_file_paths(
  path = here("data", "spectra", "2019-wizardproj"))

spc_lst_wizardproj_2019 <- map(files_wizardproj_2019, ~ future_map(.x,
  ~ read_opus_univ(fnames = .x,
    extract = c("spc")),
  .progress = TRUE)
)
```


