---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# simplerspec.opus

The goal of simplerspec.sae is to facilitate spectral measurement workflows
for Bruker spectrometers. It is tailored to file outputs and directory 
structures produced by the Bruker OPUS and OPUS Lab software.

## Installation

You can install the released version of simplerspec.opus using

``` r
devtools::install_github("philipp-baumann/simplerspec.opus")
```

## Perform folder, file, and data and metadata integrity tests

You can test the below expectations in one handy exported function:

```{r, eval = FALSE}
pkgs <- c("simplerspec.opus", "here")
lapply(pkgs, library, character.only = TRUE)

# You only need to specify the root folder of the data directory
struct_data_metadata(data_root = here("data", "spectra", "2018-BDM"))
```

This condensed integrity test verifies whether both metadata and data are
complete, for each of the dates separately.

This final check ensures that there is a minimum chance that samples were wrongly 
labelled or measurement positions were swapped.

This workflow has one main assumption:

* The binary Bruker OPUS files, that contain all spectral measurements and
  aquisition parameters, are located in folders named by date. The names
  are unluckily in the format `"YYYY_M_D"`. That does not follow 
  [ISO 8601](https://en.wikipedia.org/wiki/ISO_8601).
   
### Test category 1:

* **Expectation:** Metadata are Open Office spreadsheets with extension
  `.ods`. The file name has a date prefix of format `YYYY-MM-DD`

* **Test:** Test whether every measurement folder has a corresponding 
  metadata file. This is done by extracting `"YYYY_M_D"` from
  the metadata file names, and converting to `"YYYY-MM-DD"`, and check if
  metadata file name information and and folder names are all identical.
  
### Test category 2:

* **Expectation:** Every metadata spreadsheet has a `date` column with 
  header named identically, where entries represent dates in format
  `YYYY-MM-DD`.

* **Test:** Test whether the `date` column entries are matching the expanded
  measurement folders for each date.
  
### Tests category 3: All OPUS files have the correct metadata and vice versa

  
* **Expectations:**: The metadata spreadsheets contain a column `pos` that
  specifies the well plate positions, and also a column `sample_id`. The
  `sample_id` of a sample needs to be unique. The OPUS files that are generated 
  after measurements have a `_<plate-position>` suffix after a `sample_id`
  string.
  
* **Test:**: In order to check that there is a metadata entry for each measured
  file, file names are reconstructed from `sample_id` and `pos` information. 
  Further, repetition numbers starting from 0 are sequentially numbered by an
  integer increasing by 1, are reconstructed as file extension  
  `.<repetition_number>`. This is what is expected to be generated from the 
  measurements managed by the Bruker OPUS lab interface.
   


